---
title: "Tidymodels Tune"
author: "Chris Mann"
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(GGally)
library(finetune)
library(tidymodels)
library(tidyverse)
```

```{r}
df <- read_csv("https://raw.githubusercontent.com/marcopeix/predict-bike-rentals/master/bike_rental_hour.csv")
```
First we'll split the data to keep a small test set, taking the chronologically latest 10% of the data.
```{r}
train_test_split_date <- as.Date(quantile(unclass(df$dteday), 0.80), origin = "1970-01-01")
df <- df %>% 
  arrange(dteday)
train <- df[df$dteday <= train_test_split_date, ]
test <- df[df$dteday > train_test_split_date, ]
```

Before we start mutating the data, we'll resample the training set for cross-validation.
```{r}
slid_per <- sliding_period(train, index = dteday, period = "month")
```

We'll recode the seasons and weather as well as take the sine and cosine of `hr` to encode it numerically.

The seasons are:
1. Spring
2. Summer
3. Autumn
4. Winter

It seems likely that more people will cycle in nicer weather and less in winter.  Let's encode season based on the average count for each season in the training data.

```{r}
seasons <- train %>% 
  group_by(season) %>% 
  summarise(avg_cnt = mean(cnt)) %>% 
  mutate(levels = scale(avg_cnt)) %>% 
  select(-avg_cnt)

season_key = as.list(seasons$levels)
names(season_key) = seasons$season
season_key
```

Similarly for weather:
```{r}
weather <- train %>% 
  group_by(weathersit) %>% 
  summarise(avg_cnt = mean(cnt)) %>% 
  mutate(levels = scale(avg_cnt)) %>% 
  select(-avg_cnt)

weather_key = as.list(weather$levels)
names(weather_key) = weather$weather
weather_key
```
Now we can build out recipe.
```{r}
rec <- df %>% 
  recipe(cnt ~ .)

rec <- rec %>% 
  step_rm(casual, registered) %>% 
  step_mutate(
    season = recode(season, !!!seasons_key),
    weathersit = recode(weathersit, !!!weather_key),
    hr_sin = sin(hr),
    hr_cos = cos(hr)
    ) %>% 
  step_date(dteday) %>% 
  step_rm(dteday, hr) %>% 
  prep()
```

Let's take a look at the processed training data.
```{r}
train <- bake(rec, new_data = NULL)
train
```
```{r, warning=FALSE}
plot_data <- train %>% 
  select_if(is.numeric) %>% 
  mutate(cnt = as.factor(cnt > mean(cnt))) %>% 
  sample_n(1000)
ggscatmat(plot_data, color = "cnt", alpha = 0.4)
```


```{r}
xgb <- boost_tree(trees = 400,
                  sample_size = tune(),
                  mtry = tune(),
                  min_n = tune(),
                  tree_depth = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("regression")

params <- list(sample_size = unknown(),
               mtry = unknown(),
               min_n = unknown(),
               tree_depth = unknown())

xgb_params <- parameters(params)

update(xgb_params, mtry = finalize(mtry(), slid_per))
```

```{r}
xgb_wflow <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(xgb)
```

```{r}
xgb_grid <- xgb_wflow %>%
  parameters() %>%
  grid_max_entropy(size = 25)
```


```{r}
res <- tune_race_anova()
```

